Revealing transient strain in geodetic data with Gaussian process regression

### 2017-11-11 ResponseToReviewers/ResponseToReviewers.tex

% STRAIN COVARIANCE WAS INCORRECT

% THE NOISE MODEL WAS INCORRECT BECAUSE I USED A DIRAC RATHER THAN A
% KRONECKER DELTA.

% WE ADDED MORE POSTERIOR DISPLACEMENT FIGURES

We are grateful for the comments from you and the reviewers. We are
pleased that the comments are generally positive and that most of the
concerns seem to be with the presentation of the material rather than
the material itself. We agree that the previous draft required a
substantial amount of clarification in order to be more accessible to
geophysicists. Consequently, we have revised most of the methods
section (Section 2) with the hopes of improving clarity. Details of
our revisions are given below in our response to the reviewers'
comments.

In the revised draft, we have addressed several issues that we found,
which were not brought up by the reviewers:

1) The strain covariance function, eq. 14 of the previous draft, was
incorrect. That covariance function was only correct for the shear
strain. The normal strains differ by a factor of 2. This error was NOT
in our software implementation of the method, and so our presented
results have not changed.

2) There was an error in eq. 20 of the previous draft. We used a Dirac
delta function where we should have used a Kronecker delta.

3) In the previous draft, Figure 5 was the only figure that showed how
well our model fits the GNSS data. The data consists of seven years of
daily displacements from 94 stations, so it is not possible to show
all the data. However, we are now showing the fit at 3 stations,
rather than 1. These stations were chosen to illustrate a slight
over-smoothing in our fit to the data.

4) In addition to clarifying Section 2, we have significantly
rewritten Sections 3 (application to the Pacific Northwest) and
Section 4 (the discussion) so that they have a clearer focus and
structure. However, the content being presented has not changed.

Each of the reviewers comments are written below and followed by our
response. We hope that we have addressed their concerns sufficiently
in our revision.

@@

\textbf{Review of the paper : Revealing transient strain in geodetic data
with Gaussian process regression}

\textbf{Overview of the work: The paper focuses on a new non-parametric
Bayesian method to estimate transient strain rate from GNSS data. It
is based on Gaussian Process regression. The method also includes an
outlier removal processing step. This technique is then applied to the
detection of slow slip events in Cascadia. The validity of the
detected transient signals is accomplished via comparison with the
seismic tremor recorded along the Cascadia range.}

\textbf{The article is well-written and your algorithm is very interesting. I
think the use of Gaussian Process Regression is novel and open new
roads in GNSS data analysis. I believe that you spend a lot of efforts
in trying to justify most of the assumptions behind your model,
together with the effects of using different kind of prior. Perhaps,
that overshadows slightly your results. You could consider to
reorganize the article to make it more readable for geophysicists.}

I have taken several steps to make my revised draft more readable for
geophysicists. For details, see my response to the editor's comment
``Many of the reviewers’ comments ... ".

@@

\textbf{I also require that a comparison between your outlier removal model
and one other algorithm could help to support your claims about the
efficiency of your method.}

See my response to the comment ``Also, the paper would be improved if
the proposed outlier detection ..."

@@

  6) I provide a clearer derivation of the signal-to-noise ratio.

  7) I moved details of the outlier detection method to the appendix
     because it is not essential to the text.

  8) I expanded the explanation of the REML method, which is used to
     constrain u and eta in Section 3.1 and 3.2.

  9) I attempt to avoid esoteric jargon such as "hyperparameter". In
     the revised draft, "hyperparameter" has been replaced with
     "parameter", which does not degrade the precision of our
     description of the method.

  10) Most of the paper was reworded / restructed to improve
     clarity. Despite the fact that very little of the original text
     is in the revised daft, there are almost no differences in the
     content being presented.

@@

\textbf{2+26: possibility not risk: the latter is a technical term in seismic
hazard.}

Done

\textbf{2+47-48: I do not think this description of Shen’s method (which I’ve
used and programmed) is right, since it allows the deformation
gradients to vary in space. I’ve always thought of it as a kind of
adaptive kernel smoother, somewhat like applying 2-D loess; so it is
only depen- dent on nearby points, and has no assumption of uniformity
anywhere. (Actually this raises a more general question: is the method
given here use local or global support?)}

I have updated my description of Shen's method in the revised draft.

@@

\textbf{3+7: could also}

Done

\textbf{3+32: redo the sentence; as it stands, you are saying that the SCEC
exercise calculates strain rates.}

The text has been clarified say "The method described in Holt 2013 was
also tested in the SCEC transient detection exercise. Holt 2013
detects transient signal from an estimate of transient strain rates,
and they estimate transient strain rates using a ..."

### 2017-11-11 RevisedSubmission/hines_hetland_2017_gji.tex
handed over to eric

Displacement data from dense GNSS networks, such as the Plate Boundary
Observatory (PBO), can be used to infer secular and transient strain.
Numerous methods for converting GNSS displacement data to strain have
been proposed, and most involve assuming some parametric form for the
deformation signal. The simplest method for estimating secular strain
rates assumes that GNSS derived velocities can be described with a
first-order polynomial (i.e, the deformation gradients are constant)
over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,
Murray2000}. The components of the strain rate tensor for each
subnetwork are then determined from the least squares fit to the
observations. The assumption that deformation gradients are spatially
uniform is not appropriate when subnetworks span a very large area. To
help overcome this deficiency, \citet{Shen1996,Shen2015} used an
inverse distance weighting scheme, in which the estimated strain rate
at a point is primarily controlled by observations at nearby stations.
The method of \citet{Shen1996,Shen2015} can be viewed as a form of
local least squares regression with a first-order polynomial. Other
methods for estimating secular strain rates have parameterized GNSS
derived velocities with bi-cubic splines \citep{Beavan2001}, spherical
wavelets \citep{Tape2009}, and elastostatic Green's functions
\citep{Sandwell2016}. The type of basis functions and the number of
degrees of freedom for a parameterization, which are often chosen
subjectively, have a strong influence on the strain solution. If there
are too few degrees of freedom in the parameterization, then estimated
strain rates will be biased and the uncertainties will be
underestimated. On the other hand, if there are too many degrees of
freedom, then there will not be any coherent features in the estimated
strain rates. The methods described by \citet{Beavan2001} and
\citet{Tape2009} also require the user to specify penalty parameters
that control a similar trade-off between bias and variance in the
solution. One could also parameterize deformation with a physically
motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,
McCaffrey2007}. In such models the lithospheric rheology and fault
geometries are assumed to be known. Any errors in the assumed physical
model could result in biased strain estimates and underestimated
uncertainties.

@@

The aforementioned studies are concerned with estimating secular
strain rates. In recent years the Southern California Earthquake
Center (SCEC) community has shown interest in developing methods for
detecting transient strain. SCEC supported a transient detection
exercise \citep{Lohman2013}, where several research groups tested
their methods for detecting transient geophysical signal with a
synthetic GNSS dataset. Among the methods tested were the Network
Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion
Filter (NIF) \citep{Segall1997}. The NSF uses a wavelet
parameterization to describe the spatial component of geophysical
signal. The NIF, which is intended for imaging slow fault slip from
geodetic data, uses the elastic dislocation Green's functions from
\citet{Okada1992}. For the NSF and NIF, the time dependence of the
geophysical signal is modeled as integrated Brownian motion. The
method described in \citet{Holt2013} was also tested in the SCEC
transient detection exercise, which calculates strain rates using a
bi-cubic spatial parameterization of displacements between time
epochs. \citet{Holt2013} defined a detection threshold based on the
strain rate magnitude, and we demonstrate that this is indeed an
effective criterion for identifying a geophysical signal. For the same
reasons descibed above, the transient deformation and corresponding
uncertainties estimated by these methods can be biased by the chosen
spatial parameterization. It is then difficult to distinguish signal
from noise with these methods, which limits their utility for
transient detection.

@@

Here we propose using Gaussian process regression (GPR)
\citep{Rasmussen2006} to estimate transient strain from GNSS data. GPR
is closely related to kriging \citep{Cressie1992} and least squares
collocation \citep{Moritz1978}. The latter has been used by
\citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain
rates from GNSS data. GPR is a Bayesian, non-parametric method for
inferring a continuous function from scattered data. Since GNSS
stations are irregularly spaced and observation times may differ
between stations, GPR is an ideal tool for synthesizing discrete GNSS
data into a spatially and temporally continuous representation of
transient strain. GPR is Bayesian in that we use a prior model to
control the spatial and temporal roughness of the inferred transient
strain. The prior is specified as a stochastic process, namely a
Gaussian process. If there is no information available to help choose
an appropriate prior Gaussian process, then maximum likelihood methods
can be used to objectively choose one that is most consistent with the
data. We use GPR to infer transient strain resulting from SSEs in the
Pacific Northwest, demonstrating that GPR is an effective tool for
detecting transient geophysical processes.

@@

We let $\vec{u}(p) = (u_\e(p),u_\n(p))$ be our prior understanding of
transient displacements. Since transient displacements are not
precisely known, we cannot consider $\vec{u}$ to be a deterministic
function. Instead, $\vec{u}$ is considered to be a stochastic process
containing a distribution of functions that could potentially describe
transient displacements. Specifically, we let each component of
$\vec{u}$ be a Gaussian process. A Gaussian process is a stochastic
process whose value at any collection of points can be described with
a multivariate normal distribution. That is to say, the random vector
$[u_i(p)]_{p \in \points}$ has a multivariate normal distribution for
any collection of points $\points$. A realization of the random vector
$[u_i(p)]_{p \in \points}$ can be interpretted as a realization of
$u_i$ (i.e., a sample function) that is evaluated at $\points$. Just
as a multivariate normal distribution is fully determined by a mean
vector and a covariance matrix, a Gaussian process is fully determined
by a mean function and a covariance function. For example, Brownian
motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$
which has the mean function $\E{B(t)} = 0$ and the covariance function
$\Cov{B(t),B(t')} = \phi^2 \min(t,t')$, where $t,t' \ge 0$ and $\phi$
is a scale factor. We let each component of $\vec{u}$ have zero
mean and a generic covariance function $\Cov{u_i(p),u_i(p')} =
C_{u_i}(p,p')$. Using a more concise notation, we write our prior on
each component of $\vec{u}$ as $u_i \sim
\mathcal{GP}\left(0,C_{u_i}\right)$.

@@

We provide an example to give the prior on transient displacements a
more tangible meaning. we can use a squared exponential function for
$X_i$ and $T_i$,
\begin{equation}\label{eq:ExampleCov}
X_i(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2\ell^2}\right), \ \ \ \ 
T_i(t,t') = \phi^2 \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
\end{equation}
which satisfies our requirements for positive definiteness and
differentiability. The parameters $\ell$ and $\tau$ control the
length-scale and time-scale, respectively, of realizations of $u_i$.
The parameter $\phi$, which we have arbitrarily chosen to incorporate
into $T_i$ rather than $X_i$, controls the amplitude of realizations
of $u_i$. Ideally, we want realizations of $u_i$ to have a
length-scale, time-scale, and amplitude that resemble what we expect
for the true transient displacements. In Figure \ref{fig:PriorDemo}A,
we show $C_{u_i}$ using the squared exponential function for $X_i$ and
$T_i$ and the parameters $\ell = 100$ km, $\tau = 10$ days, and $\phi
= 1$ mm. A single realization of $u_i$ corresponsing to that choice of
covariance function is shown in Figure \ref{fig:PriorDemo}B (See
\citet[sec. A3]{Rasmussen2006} for details on drawing realization from
Gaussian processes). While the squared exponential function is a
commonly used covariance function for GPR, it is not appropriate for
every application. The appropriate choice for $X_i$ and $T_i$ may vary
depending on the geophysical signal we are trying to describe. To keep
this section sufficiently general, we hold off on specifying $X_i$ and
$T_i$ until Section \ref{sec:SignalModel}, where we estimate transient
strain from slow slip events in the Pacific Northwest.

@@

The temporally correlated noise in GNSS data has been thoroughly
studied over the past two decades (See \citet{Bock2016} or
\citet{He2017} for a review). Most studies found that the temporally
correlated noise can be described with Brownian motion (also known as
random walk noise or a Weiner process) or flicker noise
\citep[e.g.,][]{Zhang1997, Mao1999, Williams2004, Langbein2008,
Murray2017}. Less frequently used noise models include a band-pass
process \citep[e.g.,][]{Langbein2004} and a first-order Gauss-Markov
(FOGM) process \citep[e.g.,][]{Langbein2004,Dmitrieva2015}. There is
some physical justification for using Brownian motion as a noise model
because it accurately describes the power spectrum of motion resulting
from instability in a geodetic monument \citep[e.g.,][]{Wyatt1982,
Wyatt1989}. In particular, the power spectrums for Brownian motion and
monument motion both decay proportionally to $f^{-2}$, where $f$ is
frequency. However, Brownian motion necessarily contains a reference
time at which the process begins. Since there is no notion of when
noise ``begins" in GNSS data, we do not find Brownian motion to be an
appropriate noise model. On the other hand, a FOGM process has no
reference time (i.e., it is stationary), and its power spectrum,
\begin{equation}\label{eq:FOGMPS} 
S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2}, 
\end{equation} 
decays proportionally to $f^{-2}$ above the cutoff frequency $\alpha /
(2 \pi)$. We choose $z_i$ to be a spatially uncorrelated FOGM process.
The covariance function for $z_i$ is then
\begin{equation}\label{eq:FOGM}
C_{z_i}\left((\pos,t),(\pos',t')\right) = \frac{\beta^2}{2\alpha} 
                                          \exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}, 
\end{equation}
where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise
\citep[sec. B2]{Rasmussen2006}.

@@

By choosing a FOGM process for $z_i$, we have introduced two
parameters that need to be constrained, $\alpha$ and $\beta$. We
constrain these parameters, which we collectively refer to as
$\mitbf{\theta}$, with the Restricted Maximum Likelihood (REML) method
\citep{Harville1974}. The REML method is conceptually similar to the
Maximum Likelihood Estimation (MLE) method from \citet{Langbein1997},
where we pick $\mitbf{\theta}$ to maximize the probability of drawing
the observed data, $\mitbf{d}_i^*$, from the random vector
$\mitbf{d}_i$. However, unlike the MLE method, the REML method
produces unbiased estimates of $\mitbf{\theta}$ \citep[sec.
2.6]{Cressie1992}. We use the REML method to estimate $\mitbf{\theta}$
at 38 continuous GNSS stations in the Pacific Northwest that are east
of 121$^\circ$W. These stations are sufficiently far from the
subduction zone that they are unlikely to record transient deformation
from SSEs, allowing us to ingore the term $u_i(\points)$ in
$\mitbf{d}_i$. We assume the noise at these inland stations is
representative of the noise at all the stations considered in this
study, which is probably a poor assumption since the inland stations
are subject to distinctly different climatic conditions. Nonetheless,
we find $\mitbf{\theta}$ for each of the inland stations and for each
displacement component by maximizing the REML likelihood function
\begin{equation}\label{eq:REMLNoise}
\mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\G^T\G\right|}{(2\pi)^{n-6m} 
                                    \left| \mitbf{C}_{\eta_i}(\mitbf{\theta}) \right| 
                                    \left| \G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G \right|}\right)^{\frac{1}{2}} 
                              e^{-\tfrac{1}{2}\data_i^{*T}\mitbf{K}(\mitbf{\theta})\data_i^*}
\end{equation}
with respect to $\mitbf{\theta}$, where
\begin{equation}
\mitbf{K}(\mitbf{\theta}) = \mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1} - 
                            \mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G
                            \left(\G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G\right)^{-1}
                            \G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}.
\end{equation}
In the above equation, $\data^*_i$ consists of the data for a single station.

@@

In addition to detecting geophysical processes, the GNSS derived
transient strain rates can be used to better understand the data from
borehole strain meters (BSMs). The Plate Boundary Observatory contains
about forty BSMs in the Pacific Northwest, and it has been
demonstrated that BSMs are able to record transient geophysical events
such as SSEs \citep[e.g.,][]{Dragert2011}. However, there are
complications that prevent BSM data from being used quantitatively in
geophysical studies. One difficulty is that BSM data should be
calibrated with a well known strain source, such as diurnal and
semidiurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
Unfortunately, the tidal forces at BSMs which record SSEs are strongly
influenced by local bodies of water such as the Straight of Juan de
Fuca, making it difficult to form a theoretical prediction of tidal
strains \citep{Roeloffs2010}. Another complication is that noise in
BSM data is not well understood. The noise consists, in part, of a
long-term decay resulting from the instrument equilibrating with the
surrounding rock \citep{Gladwin1987}. Typically, this noise is dealt
with in an ad-hoc manner by fitting and removing exponentials and
low-order polynomials. We envision that the GNSS derived strain rates
from this paper can be used as a reference strain for calibrating BSM
data and quantify its noise.

>>>>

Displacement data from dense GNSS networks, such as the Plate Boundary
Observatory (PBO), can be used to infer secular and transient strain.
Numerous methods for converting GNSS displacement data to strain have
been proposed, and most involve assuming some parametric form for the
deformation signal. The simplest method for estimating secular strain
rates assumes that GNSS derived velocities can be described with a
first-order polynomial (i.e, the deformation gradients are constant)
over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,
Murray2000}. The components of the secular strain rate tensor for each
subnetwork are then determined from the least squares fit to the
observations. The assumption that deformation gradients are spatially
uniform is not appropriate when subnetworks span a very large area. To
help overcome this deficiency, \citet{Shen1996,Shen2015} used an
inverse distance weighting scheme, in which the estimated secular
strain rate at a point is primarily controlled by observations at
nearby stations. The method of \citet{Shen1996,Shen2015} can be viewed
as a form of local least squares regression with a first-order
polynomial. Other methods for estimating secular strain rates have
parameterized GNSS derived velocities with bi-cubic splines
\citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and
elastostatic Green's functions \citep{Sandwell2016}. The type of basis
functions and the number of degrees of freedom for a parameterization,
which are often chosen subjectively, have a strong influence on the
solution. If there are too few degrees of freedom in the
parameterization, then estimated secular strain rates will be biased
and the uncertainties will be underestimated. On the other hand, if
there are too many degrees of freedom, then there will not be any
coherent features in the estimated secular strain rates. The methods
described by \citet{Beavan2001} and \citet{Tape2009} also require the
user to specify penalty parameters that control a similar trade-off
between bias and variance in the solution. One could also parameterize
deformation with a physically motivated model of interseismic
deformation \citep[e.g.,][]{Meade2005, McCaffrey2007}. In such models
the lithospheric rheology and fault geometries are assumed to be
known. Any errors in the assumed physical model could result in biased
estimates of the secular strain rates and underestimated
uncertainties.

@@

The aforementioned studies are concerned with estimating secular
strain rates. In recent years the Southern California Earthquake
Center (SCEC) community has shown interest in developing methods for
detecting transient deformation. SCEC supported a transient detection
exercise \citep{Lohman2013}, where several research groups tested
their methods for detecting transient geophysical signal with a
synthetic GNSS dataset. Among the methods tested were the Network
Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion
Filter (NIF) \citep{Segall1997}. The NSF uses a wavelet
parameterization to describe the spatial component of geophysical
signal. The NIF, which is intended for imaging slow fault slip from
geodetic data, uses the elastic dislocation Green's functions from
\citet{Okada1992}. For the NSF and NIF, the time dependence of the
geophysical signal is modeled as integrated Brownian motion. The
method described in \citet{Holt2013} was also tested in the SCEC
transient detection exercise. \citet{Holt2013} detect transient
signal from an estimate of strain rates, and they estimate strain
rates using a bi-cubic spatial parameterization of displacements
between time epochs. \citet{Holt2013} define a detection threshold
based on the strain rate magnitude, and we demonstrate that this is
indeed an effective criterion for identifying a geophysical signal.
For the same reasons described above, the transient deformation and
corresponding uncertainties estimated by these methods can be biased
by the chosen spatial parameterization. It is then difficult to
distinguish signal from noise with these methods, which limits their
utility for transient detection.

@@

Here we propose using Gaussian process regression (GPR)
\citep{Rasmussen2006} to estimate transient strain rates from GNSS
data. GPR is closely related to kriging \citep{Cressie1992} and least
squares collocation \citep{Moritz1978}. The latter has been used by
\citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain
rates from GNSS data. GPR is a Bayesian, non-parametric method for
inferring a continuous function from scattered data. Since GNSS
stations are irregularly spaced and observation times may differ
between stations, GPR is an ideal tool for synthesizing discrete GNSS
data into a spatially and temporally continuous representation of
transient strain rates. GPR is Bayesian in that we use a prior model
to control the spatial and temporal roughness of the inferred
transient strain rates. The prior is specified as a stochastic
process, namely a Gaussian process. If there is no information
available to help choose an appropriate prior Gaussian process, then
maximum likelihood methods can be used to objectively choose one that
is most consistent with the data. We use GPR to infer transient strain
rates resulting from SSEs in the Pacific Northwest, demonstrating that
GPR is an effective tool for detecting transient geophysical
processes.

@@

We let $\vec{u}(p) = (u_\e(p),u_\n(p))$ be our prior understanding of
transient displacements. Since transient displacements are not
precisely known, we cannot consider $\vec{u}$ to be a deterministic
function. Instead, $\vec{u}$ is considered to be a stochastic process
containing a distribution of functions that could potentially describe
transient displacements. Specifically, we let each component of
$\vec{u}$ be a Gaussian process. A Gaussian process is a stochastic
process whose value at any collection of points can be described with
a multivariate normal distribution. That is to say, the random vector
$[u_i(p)]_{p \in \points}$ has a multivariate normal distribution for
any collection of points $\points$. A realization of the random vector
$[u_i(p)]_{p \in \points}$ can be interpreted as a realization of
$u_i$ (i.e., a sample function) that is evaluated at $\points$. Just
as a multivariate normal distribution is fully determined by a mean
vector and a covariance matrix, a Gaussian process is fully determined
by a mean function and a covariance function. For example, Brownian
motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$
which has the mean function $\E{B(t)} = 0$ and the covariance function
$\Cov{B(t),B(t')} = \phi^2 \min(t,t')$, where $t,t' \ge 0$ and $\phi$
is a scale factor. We let each component of $\vec{u}$ have zero
mean and a generic covariance function $\Cov{u_i(p),u_i(p')} =
C_{u_i}(p,p')$. Using a more concise notation, we write our prior on
each component of $\vec{u}$ as $u_i \sim
\mathcal{GP}\left(0,C_{u_i}\right)$.

@@

We provide an example to give the prior on transient displacements a
more tangible meaning. we can use a squared exponential function for
$X_i$ and $T_i$,
\begin{equation}\label{eq:ExampleCov}
X_i(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2\ell^2}\right), \ \ \ \ 
T_i(t,t') = \phi^2 \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
\end{equation}
which satisfies our requirements for positive definiteness and
differentiability. The parameters $\ell$ and $\tau$ control the
length-scale and time-scale, respectively, of realizations of $u_i$.
The parameter $\phi$, which we have arbitrarily chosen to incorporate
into $T_i$ rather than $X_i$, controls the amplitude of realizations
of $u_i$. Ideally, we want realizations of $u_i$ to have a
length-scale, time-scale, and amplitude that resemble what we expect
for the true transient displacements. In Figure \ref{fig:PriorDemo}A,
we show $C_{u_i}$ using the squared exponential function for $X_i$ and
$T_i$ and the parameters $\ell = 100$ km, $\tau = 10$ days, and $\phi
= 1$ mm. A single realization of $u_i$ corresponding to that choice of
covariance function is shown in Figure \ref{fig:PriorDemo}B (See
\citet[sec. A3]{Rasmussen2006} for details on drawing realization from
Gaussian processes). While the squared exponential function is a
commonly used covariance function for GPR, it is not appropriate for
every application. The appropriate choice for $X_i$ and $T_i$ may vary
depending on the geophysical signal we are trying to describe. To keep
this section sufficiently general, we hold off on specifying $X_i$ and
$T_i$ until Section \ref{sec:SignalModel}, where we estimate transient
strain from slow slip events in the Pacific Northwest.

@@

The temporally correlated noise in GNSS data has been thoroughly
studied over the past two decades (See \citet{Bock2016} or
\citet{He2017} for a review). Most studies found that the temporally
correlated noise can be described with Brownian motion (also known as
random walk noise or a Weiner process) or flicker noise
\citep[e.g.,][]{Zhang1997, Mao1999, Williams2004, Langbein2008,
Murray2017}. Less frequently used noise models include a band-pass
process \citep[e.g.,][]{Langbein2004} and a first-order Gauss-Markov
(FOGM) process \citep[e.g.,][]{Langbein2004,Dmitrieva2015}. There is
some physical justification for using Brownian motion as a noise model
because it accurately describes the power spectrum of motion resulting
from instability in a geodetic monument \citep[e.g.,][]{Wyatt1982,
Wyatt1989}. In particular, the power spectra for Brownian motion and
monument motion both decay proportionally to $f^{-2}$, where $f$ is
frequency. However, Brownian motion necessarily contains a reference
time at which the process begins. Since there is no notion of when
noise ``begins" in GNSS data, we do not find Brownian motion to be an
appropriate noise model. On the other hand, a FOGM process has no
reference time (i.e., it is stationary), and its power spectrum,
\begin{equation}\label{eq:FOGMPS} 
S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2}, 
\end{equation} 
decays proportionally to $f^{-2}$ above the cutoff frequency $\alpha /
(2 \pi)$. We choose $z_i$ to be a spatially uncorrelated FOGM process.
The covariance function for $z_i$ is then
\begin{equation}\label{eq:FOGM}
C_{z_i}\left((\pos,t),(\pos',t')\right) = \frac{\beta^2}{2\alpha} 
                                          \exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}, 
\end{equation}
where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise
\citep[sec. B2]{Rasmussen2006}.

@@

By choosing a FOGM process for $z_i$, we have introduced two
parameters that need to be constrained, $\alpha$ and $\beta$. We
constrain these parameters, which we collectively refer to as
$\mitbf{\theta}$, with the Restricted Maximum Likelihood (REML) method
\citep{Harville1974}. The REML method is conceptually similar to the
Maximum Likelihood Estimation (MLE) method from \citet{Langbein1997},
where we pick $\mitbf{\theta}$ to maximize the probability of drawing
the observed data, $\mitbf{d}_i^*$, from the random vector
$\mitbf{d}_i$. However, unlike the MLE method, the REML method
produces unbiased estimates of $\mitbf{\theta}$ \citep[sec.
2.6]{Cressie1992}. We use the REML method to estimate $\mitbf{\theta}$
at 38 continuous GNSS stations in the Pacific Northwest that are east
of 121$^\circ$W. These stations are sufficiently far from the
subduction zone that they are unlikely to record transient deformation
from SSEs, allowing us to ignore the term $u_i(\points)$ in
$\mitbf{d}_i$. We assume the noise at these inland stations is
representative of the noise at all the stations considered in this
study, which is probably a poor assumption since the inland stations
are subject to distinctly different climatic conditions. Nonetheless,
we find $\mitbf{\theta}$ for each of the inland stations and for each
displacement component by maximizing the REML likelihood function
\begin{equation}\label{eq:REMLNoise}
\mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\G^T\G\right|}{(2\pi)^{n-6m} 
                                    \left| \mitbf{C}_{\eta_i}(\mitbf{\theta}) \right| 
                                    \left| \G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G \right|}\right)^{\frac{1}{2}} 
                              e^{-\tfrac{1}{2}\data_i^{*T}\mitbf{K}(\mitbf{\theta})\data_i^*}
\end{equation}
with respect to $\mitbf{\theta}$, where
\begin{equation}
\mitbf{K}(\mitbf{\theta}) = \mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1} - 
                            \mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G
                            \left(\G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G\right)^{-1}
                            \G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}.
\end{equation}
In the above equation, $\data^*_i$ consists of the data for a single station.

@@

In addition to detecting geophysical processes, the GNSS derived
transient strain rates can be used to better understand the data from
borehole strain meters (BSMs). The Plate Boundary Observatory contains
about forty BSMs in the Pacific Northwest, and it has been
demonstrated that BSMs are able to record transient geophysical events
such as SSEs \citep[e.g.,][]{Dragert2011}. However, there are
complications that prevent BSM data from being used quantitatively in
geophysical studies. One difficulty is that BSM data should be
calibrated with a well known strain source, such as diurnal and
semi-diurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
Unfortunately, the tidal forces at BSMs which record SSEs are strongly
influenced by local bodies of water such as the Straight of Juan de
Fuca, making it difficult to form a theoretical prediction of tidal
strains \citep{Roeloffs2010}. Another complication is that noise in
BSM data is not well understood. The noise consists, in part, of a
long-term decay resulting from the instrument equilibrating with the
surrounding rock \citep{Gladwin1987}. Typically, this noise is dealt
with in an ad-hoc manner by fitting and removing exponentials and
low-order polynomials. We envision that the GNSS derived strain rates
from this paper can be used as a reference strain for calibrating BSM
data and quantify its noise.
